MATH 245 Homework 2

### 2024-02-19 PDEs/math245_hw2.Rmd
pde hw revisions

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-17"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$

@@

@@

First, we can factor the equation as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

@@

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
\end{tcolorbox}

@@

\begin{equation} 
\Phi_{\text{even}}=
    \begin{cases} 
      \Phi(x) & \text{for } x \geq 0\\
      \Phi(-x) & \text{for } x \leq 0\\
    \end{cases}  
    ,\quad
\Psi_{\text{even}}=
    \begin{cases}
      \Psi(x) & \text{for } x \geq 0\\
      \Psi(-x) & \text{for } x \leq 0\\
    \end{cases}
    ,\quad
F_{\text{even}}=
    \begin{cases} 
      F(t,x) & \text{for } x \geq 0\\
      F(t,-x) & \text{for } x \leq 0\\
      \end{cases}
\end{equation}

@@

\begin{equation*}
\frac{\phi(x_0+ct_0)-H(0)+\phi(x_0-ct_0)-H(0)}{2} +
\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x)-H'(0)dx + \\
\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x)-H''(t) dx\,dt
\end{equation*}

\begin{align*}
& w(t_0,x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2}- H(0) +
\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx - \frac{H'(0)}{2c}\left[(x_0+ct_0)-(x_0-ct_0)\right]+ \\
& \frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt -
\frac{1}{2c} H''(t) \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\,dt
\end{align*}

@@

Now we use integration by parts on the final integral in (\ref{eqn:w1}), setting
$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$,

@@

@@

$$\text{III}= \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\;dt $$
$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;2c(t_0-t)\;dt $$
$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;(t_0-t)\;dt $$
Now we use integration by parts on the final integral in the above expression, setting
$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$

@@

$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right),$$
which simplifies to
$$w_{tt}=c^2 w_{rr}.$$
But this is just the wave equation, and we can use d'Alembert's formula to find the solution:
 $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
 Since $\varphi=0$, 
 $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$

@@

\begin{equation}\label{eqn:prob5c}
=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{2u_x u_{xt}}\:dx =: I+J.
\end{equation}

@@

\begin{equation} \label{eqn:J}
J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_xx u_t}\:dx + 
\left[ u_x u_t \right]_{x=0}^{x=1}.
\end{equation}

@@

Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(z_1)=g\left(\frac{z_1}{2}\right)-h_2(0)$.

Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(z_2)=f\left(\frac{z_2}{2}+2\right)-h_1(4)$.

Thus,
$$h_1(z_1)+h_2(z_2)=g\left(\frac{z_1}{2}\right)-h_2(0)+f\left(\frac{z_2}{2} + 2\right)-h_1(4)$$

@@

so after replacing $z_1$ by $x+\frac{y^2}{2}$ and $z_2$ by $x-\frac{y^2}{2}$, we get

@@

$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f'(2)$$

$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f''(2)$$

>>>>

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-18"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

a) $$u_{xx}-y^2u_{yy}+u_x-u+x^2=0 $$

@@

Both PDEs are nowhere singular, because the coefficient of $u_{xx}$ is never zero.

@@

We are looking for a solution $u(x,y)$ for the above IVP. First, we can factor the given PDE as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

@@

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}$$
\end{tcolorbox}

@@

\begin{equation} 
\Phi_{\text{even}}(x)=
    \begin{cases} 
      \Phi(x) & \text{for } x \geq 0\\
      \Phi(-x) & \text{for } x \leq 0\\
    \end{cases}  
    ,\quad
\Psi_{\text{even}}(x)=
    \begin{cases}
      \Psi(x) & \text{for } x \geq 0\\
      \Psi(-x) & \text{for } x \leq 0\\
    \end{cases}
    ,\quad
F_{\text{even}}(t,x)=
    \begin{cases} 
      F(t,x) & \text{for } x \geq 0\\
      F(t,-x) & \text{for } x \leq 0\\
      \end{cases}
\end{equation}

@@

\begin{equation*}
=\frac{\phi(x_0+ct_0)-H(0)+\phi(x_0-ct_0)-H(0)}{2} +
\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x)-H'(0)dx + \\
\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x)-H''(t) dx\,dt
\end{equation*}

\begin{align*}
& w(t_0,x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2}- H(0) +
\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx - \frac{H'(0)}{2c}\left[(x_0+ct_0)-(x_0-ct_0)\right]+ \\
& \frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt -
\frac{1}{2c} \int_{0}^{t_0} H''(t)  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\,dt
\end{align*}

@@

Now we use integration by parts on the final integral in (\ref{eqn:w1}), setting
$$u=(t_0-t), \quad du=-dt, \quad dv=H''(t)dt, \quad v=H'(t)$$,

@@

Where $\Delta$ is the 2D  region in the diagram below.

@@

$$\text{III}= \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\;dt $$
$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;2c(t_0-t)\;dt $$
$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;(t_0-t)\;dt $$
Now we use integration by parts on the final integral in the above expression, setting
$$u=(t_0-t), \quad du=-dt, \quad dv=H''(t)dt, \quad v=H'(t)$$

@@

$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right),$$
which simplifies to
$$w_{tt}=c^2 w_{rr}.$$
But this is just the wave equation, giving us the following IVP:

\begin{equation*} 
\begin{cases}
w_{tt}=c^2 w_{rr} \\
u(0,r)=0 \\
u_t(0,r) = \begin{cases}
  1  & \lvert{r}\rvert \leq 1 \\
  0 &  \lvert{r}\rvert > 1
  \end{cases}
\end{cases}
\end{equation*}

So can use d'Alembert's formula to find the solution:
 $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
 Since $\varphi=0$, 
 $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$

@@

\begin{equation}\label{eqn:prob5c}
=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{u_x u_{xt}}\:dx =: I+J.
\end{equation}

@@

\begin{equation} \label{eqn:J}
J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_{xx} u_t}\:dx + 
\left[ u_x u_t \right]_{x=0}^{x=1}.
\end{equation}

@@

Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(x)=g\left(\frac{x}{2}\right)-h_2(0)$.

Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(x)=f\left(\frac{x}{2}+2\right)-h_1(4)$.

Thus,
$$h_1(x)+h_2(x)=g\left(\frac{x}{2}\right)-h_2(0)+f\left(\frac{x}{2} + 2\right)-h_1(4)$$

@@

so after replacing the $x$ in $h_1(x)$ by $x+\frac{y^2}{2}$ and the $x$ in $h_2(x)$ by $x-\frac{y^2}{2}$, we get

@@

$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$

$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$
